---
title: "HW2_ahcooper"
author: "Andrew Cooper"
date: "9/15/2020"
output: pdf_document
---

```{r echo = F, message = F, warning = F}
library(readr)
library(tidyverse)
library(knitr)
```


## Problem 3

I aim to make version control an integral and necessary component of my future career as a statistician/data analyst. After programming for about 6 years I have realized how much time I have dedicated to sorting through old versions of code or struggling to find the most up-to-date version of a project. I have also become aware of how often I duplicate the same work for multiple projects. I plan on using version control to keep a continuously up-to-date system of all my projects so that I can always grab and use past code whenever I need to.

# Problem 4

## Part a

### base R data-munging

```{r message = F, warning = F}
# Read in raw sensory data
Sensory_raw <- read_table2("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat", 
     col_types = cols(Item = col_character()), col_names = F, skip = 2)
kable(Sensory_raw[1:5,])
```

We see the issue with the raw data is that measurements for a given item are on multiple rows. To fix this, I take each row of measurements, re-align it such that no columns are empty, and add a new column listing the correct item for each row.

```{r}
Sensory_values <- apply(Sensory_raw, 1, function(x){
  if(any(is.na(x))){
    return(x[!is.na(x)])
  }else{
    return(x[-1])
  }
})
Sensory <- data.frame(cbind(rep(seq(1, 10, 1), each = 3), t(Sensory_values)))
colnames(Sensory) <- c("Item", seq(1, 5, 1))
Sensory <- as_data_frame(Sensory)
```

### tidyverse data-munging

The measurements for each operator are currently spread out over multiple columns. A cleaner version would have these measurements collected into a single column, which I do using the "gather" function from tidyverse.

```{r}
Sensory %>% 
  gather(Operator, Value, -Item) %>% 
  group_by(Operator) %>% 
  summarize(`Average Value` = mean(Value)) %>% 
  kable()
```

### Plot

Given the nature of the data I decided to create a bar-chart in ggplot, where my summary statistic is the average measurement across all items for a single operator.


```{r echo = F}
Sensory %>% 
  mutate(Item = as.factor(Item)) %>% 
  gather(Operator, Value, -Item) %>% 
  group_by(Operator) %>% 
  summarize(Avg_Value = mean(Value)) %>% 
  ggplot(aes(x = Operator, y = Avg_Value)) + 
  geom_col() + 
  labs(y = "Average Value", title = "Average Measurements for Five Operators Across Ten Items")
```


## Part b

### base R data-munging

```{r message = F, warning = F}
Medals_raw <- read_table2("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat")
kable(Medals_raw[1:5,])
```

We see the problem with this raw table is the data aren't organized into columns, but follow the pattern of year and maximum long jump. To fix this, I extract the raw data and put it into a vector, then use the pattern to organize the data into two appropriate columns.


```{r message = F, warning = F}
Medals_values <- as.numeric(c(t(Medals_raw)))
Medals_values <- Medals_values[!is.na(Medals_values)]
Medals <- data.frame(Year = Medals_values[seq(1, length(Medals_values), 2)],
                     Long_Jump = Medals_values[seq(2, length(Medals_values), 2)])
Medals <- as_data_frame(Medals)
```


### tidyverse data-munging

We also need to mutate the year variable since it is currently coded. The mutation is simple; we just add 1990 to each value. We also sort the rows by year, which makes sense intuitively.

```{r}
Medals <- Medals %>% 
  mutate(Year = Year + 1990) %>% 
  arrange(Year)
```

```{r echo = F}
kable(summary(Medals %>% select(Year)))
```


```{r echo = F}
kable(summary(Medals %>% select(Long_Jump)))
```

### Plot

A line-plot of the long jump data seems appropriate given the time-series nature of the data.

```{r echo = F}
Medals %>% as_data_frame() %>% 
  arrange(Year) %>% 
  ggplot(aes(x = Year, y = Long_Jump)) + 
  geom_line() + 
  labs(y = "Long Jump (cm)")
```


## Part c

### base R data-munging

```{r message = F, warning = F}
Body_raw <- read_table2("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat")
kable(Body_raw[1:5,])
```

We see the problem with the raw data is similar to the problem in part b; that is, the data isn't organized into proper columns but uses a pattern of "body weight", "brain weight" for each observation. Like in part b, I extracted the raw data and put it into a vector, then used the pattern to organize the data into two appropriate columns.

```{r message = F, warning = F}
Body_values <- as.numeric(c(t(Body_raw)))
Body_values <- Body_values[!is.na(Body_values)]
Body <- data.frame(Brain_Weight = Body_values[seq(1, length(Body_values), 2)],
                     Body_Weight = Body_values[seq(2, length(Body_values), 2)])
Body <- as_data_frame(Body)
```

### tidyverse data-munging

```{r}
Body %>% 
  summarize(`Average Body Weight` = mean(Body_Weight),
            `Average Brain Weight` = mean(Brain_Weight),
            `Body Weight SD` = sd(Body_Weight),
            `Brain Weight SD` = sd(Brain_Weight),
            `Min Body Weight` = min(Body_Weight),
            `Min Brain Weight` = min(Body_Weight),
            `Max Body Weight` = max(Body_Weight),
            `Max Brain Weight` = max(Body_Weight)) %>% 
  kable()
```


### Plot

Due to the extremely positive-skewed brain and body weights as we see in the summary table above, it seems appropriate to log-transform both variables in the x-y plot.

```{r echo = F}
Body %>% 
  mutate(log_brain = log(Brain_Weight),
         log_body = log(Body_Weight)) %>% 
  ggplot(aes(x = log_body, y = log_brain)) + 
  geom_point() + 
  labs(x = "log-Body Weight", y = "log-Body Weight", title = "Body Weights and Brain Weights of Various Animals")
```


We observe a strong positive linear relationship between log-body weight and log-brain weight.


## Part d

### base r data-munging

```{r, warning = F, message = F}
Tomatoes_raw <- read_table2("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat", skip = 2, 
                   col_names = c("Tomato_Type", "1000", "2000", "3000"))
kable(Tomatoes_raw[1:5,])
```

We see the problem with the raw table is that multiple observations are within columns in a comma-delimited format. To fix this, I use apply to iterate through each row and split up these values accordingly.

```{r warning = F, message = F}
Tomatoes <- data.frame(Tomato_Type = rep(Tomatoes_raw$Tomato_Type, each = 3), 
                       apply(Tomatoes_raw[,-1], 2, function(x){as.numeric(unlist(strsplit(x, ",")))}))
Tomatoes <- as_data_frame(Tomatoes)
```

### tidyverse data-munging

Currently all the yield measurements for different densities are spread out across multiple columns. It would be tidier to gather these into a single column, which I do in tidyverse.


```{r}
Tomatoes <- Tomatoes %>% 
  gather(Density, Yield, -Tomato_Type)
```


```{r echo = F}
Tomatoes %>% 
  group_by(Tomato_Type, Density) %>% 
  summarize(Avg_Yield = mean(Yield)) %>% 
  kable()
```


### Plot

Barplots again seem appropriate here to show the average measured yield for each density and tomato type.

```{r echo = F, warning = F, message = F}
Tomatoes %>% 
  group_by(Tomato_Type, Density) %>% 
  summarize(Avg_Yield = mean(Yield)) %>% 
  ggplot(aes(x = Density, y = Avg_Yield, fill = Tomato_Type)) + 
  geom_col(position = "dodge") + 
  labs(y = "Average Yield")
```

